Reproducible Research, Peer Assessment #1
Coursera, Johns Hopkins, Prof. Peng

========================================================
## Loading and preprocessing the data
### Load libraries
```{r packages_used, echo=T}
library('ggplot2')
library('RColorBrewer')
library('lubridate')
```
### Acquire Data
```{r fetch_data, echo=TRUE}
sourceUrl = 'https://d396qusza40orc.cloudfront.net'
sourceFileName = 'repdata_data_activity.zip'
zipFile = "activity.zip"
csvFile = 'activity.csv'
### readData reads in datafile if it exists, or recreates it as much as necessary
###          if downloading, notes the download date in file 'dateUrlDownloaded.txt'
readData = function() {
  if(!file.exists(csvFile)) { # if the csvFile exists, move on
    		if(!file.exists(zipFile)) { # only download when no zipFile in dir
        		download.file(sourceUrl, destfile=zipFileName, method="curl")
			sink('./dateUrlDownloaded.txt')
			downloadDate = date()   
			sink()
		} #end download sourceUrl
		unzip(zipFile)
	} 
    	dataIn = read.csv(csvFile, header=T, sep=',') 
} 

```
### Load and Examine Data
```{r examine_data, echo=T}
activityData = readData()
names(activityData)
str(activityData)
summary(activityData)
dim(activityData)
summary(activityData$steps)
```
# What is the mean total number of steps taken per day?
```{r total_steps, echo=T}
dailySteps <- aggregate(steps ~ date, data=activityData, FUN=sum)
meanDailySteps = mean(dailySteps$steps)
medianDailySteps = median(dailySteps$steps)
maxStepsPerDay = dailySteps$day[which.max(dailySteps$steps)]
intervalSteps <- aggregate(steps ~ interval, data=activityData, FUN=sum)
meanIntervalSteps = mean(intervalSteps$steps)
medianIntervalSteps = median(intervalSteps$steps)
maxStepsPerInterval = intervalSteps$interval[which.max(intervalSteps$steps)]
```

## What is the average daily activity pattern?
```{r ave_activity, echo=T}
### plot number of steps per day
barplot(dailySteps$steps, names.arg=dailySteps$date, xlab="Date", ylab="Steps",cex.names=0.5,col='blue')

### plot number of steps per time interval
barplot(intervalSteps$steps, names.arg=intervalSteps$interval, col='blue', xlab="Time Interval", ylab="Steps", cex.names=0.5, main='Steps per Interval over the Course of a Day')

```
## Imputing missing values
```{r spline_NAs, echo=T}
sum(is.na(activityData))
sum(is.na(activityData$steps))
sum(is.na(activityData$date))
sum(is.na(activityData$interval))
missingVals = is.na(activityData$steps)
str(missingVals)

activityData <- merge(activityData, intervalSteps, by="interval", suffixes=c("",".splined"))
names(activityData)
activityData$steps[missingVals] <- activityData$steps.splined[missingVals]
splinedDailySteps <- aggregate(steps.splined ~ date, data=activityData, FUN=sum)
sum(is.na(splinedDailySteps))
meanSplinedDailySteps = mean(splinedDailySteps$steps)
medianSplinedDalySteps = median(splinedDailySteps$steps)
```
## Are there differences in activity patterns between weekdays and weekends?
```{r weekends_different, echo=T}
is.weekend = function(date){
  if (wday(date, label=T) %in% c("Sat", "Sun")){ 
    "Weekend"}
  else {
    "Weekday"}
}
activityData$isWeekend = as.factor(sapply(activityData$date, is.weekend))
names(activityData)

weekendData = subset(activityData, isWeekend == 'Weekend')
weekdayData = subset(activityData, isWeekend == 'Weekday')
par(mfrow=c(2,1))
plot(weekendData$interval, weekendData$steps, type='h', xlab="Time Interval", ylab="Steps per Interval", col='blue', main='Weekend Data')
plot(weekdayData$interval, weekdayData$steps, type='h', xlab="Time Interval", ylab="Steps per Interval", col='red', main='Weekday Data')
```

## Report Date
```{r document_report_date, echo=T}
### Document report date
reportDate = date()
```

